Using Scipion for Stream Image Processing at Cryo-EM Facilities

### 2018-04-25 2.overall.tex
final hopefully

The workflow starts two independent protocols in parallel. The first protocol estimates the camera gain at each pixel for a few  movies \citep{sorzano2018}. % In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1. 
The output of this protocol serves as verification of the validity of the experimental gain image for the whole data set. The second protocol performs direct detector movie alignment. Currently \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}. Then the CTF of the resulting micrographs are calculated. Scipion provides three CTF estimation methods:
CTFFIND4 (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013). At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that reports up to which frequency two estimations of the same micrograph are equivalent. Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus. %Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.

>>>>

The workflow starts two independent protocols in parallel. The first protocol estimates the camera gain at each pixel for a few  movies \citep{sorzano2018}. % In the CNB setup the microscope camera is calibrated and internally performs the gain correction before writing the movies so the estimated gain image should be almost constant and equal to 1. 
The output of this protocol serves as verification of the validity of the experimental gain image for the whole data set. The second protocol performs direct detector movie alignment. Currently, \scipion offers  five  methods for movie alignment: motioncorr \citep{Li2015}, motioncor2 \citep{Zheng2016:motioncor2}, correlation (an unpublished CPU version that implements in Xmipp a global alignment method and solves it similarly to motioncorr but using robust statistics), unblur \citep{Grant2015:unblur} and optical flow \citep{Abrishami2015:optical-flow}. Then the CTFs of the resulting micrographs are calculated. Scipion provides three CTF estimation methods:
CTFFIND4 (Rohou and Grigorieff, 2015), gCTF (Zhang, 2016) and Xmipp CTF estimation (Sorzano et al., 2007; Vargas et al., 2013). At this point the user may compute the CTFs using several algorithms and compare them with the protocol \protocol{CTF consensus} that reports up to which frequency two estimations of the same micrograph are equivalent. Further automatic filtering based on CTF is available through the protocol \protocol{CTF filtering} that selects micrographs based on the astigmatism, resolution and defocus. %Another criteria for CTF selection is iciness, that is,  the ratio of the intensity in the resolution band between 3.5 and 3.9 \AA, and between 6 and 30 \AA, which gives a good estimate of the ice crystal content of the vitreous specimen.

### 2018-04-25 3.workflow_cnb.tex
final hopefully

At the beginning CNB networking setup was a quite straightforward one, movies where stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine. The situation changed when the new Falcon-III detector was installed. After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory. In the following we describe such a pipeline.

@@

The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).

Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope  turn-shift starts. When two, or even three, users try to download their
data simultaneously, output-NAS works as front-end and isolates the data
acquiring and processing infrastructure from the user’s downloading activities.
Network users access to output-NAS using a hardened common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. Weak ciphers (as arcfour) has been activated in outputNAS to speed up transfer.

@@

\begin{enumerate}
 \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files and use them to create new predetermined workflows.
 \item  Automatic creation of pdf user’s report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
 \item The above mentioned information is stored in a database for further analysis. By default, the system plots the  average resolution and astigmatism of the different acquisitions versus time.
 \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion.
 \item Automatic backup of the \scipion project.
 \item \scipion is able to create live HTML reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information.
\end{enumerate}

>>>>

At the beginning CNB networking setup was a quite straightforward one, movies were stored in the  same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine. The situation changed when the new Falcon-III detector was installed. After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory. In the following we describe such a pipeline.

@@

The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).

Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope  turn-shift starts. When two, or even three, users try to download their
data simultaneously, output-NAS works as front-end and isolates the data
acquiring and processing infrastructure from the user's downloading activities.
Network users access to output-NAS using a hardened common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. Weak ciphers (as arcfour) has been activated in outputNAS to speed up transfer.

@@

\begin{enumerate}
 \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this file and use them to create new predetermined workflows.
 \item  Automatic creation of pdf user's report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
 \item The above mentioned information is stored in a database for further analysis. By default, the system plots the  average resolution and astigmatism of the different acquisitions versus time.
 \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion.
 \item Automatic backup of the \scipion project.
 \item \scipion is able to create live HTML reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information.
\end{enumerate}

### 2018-04-25 3.workflow_diamond.tex
final hopefully

Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential. A user management system
has been developed to allocate each user with a unique federal identity (ID),
and to associate this ID with the experiments in which the user partake in.
 Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory. Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting. %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF. Delageniere et al Bioinformatics 2011). In contrast to the CNB setup, Diamond has integrated the microscopes in much the same fashion to that of the beamlines at the facility, as is shown in figure X

>>>>

Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential. A user management system
has been developed to allocate each user with a unique federal identity (ID),
and to associate this ID with the experiments in which the user partakes in.
 Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory. Users may access their data through SynchWeb which fronts ISPyB, a Laboratory Information Management System combining sample tracking and experiment reporting \citep{Delageniere2011:ispb, Fisher2015:synchweb}.

 %Users may access their data through ISPyB \citep{Delageniere2011:ispb},  a Laboratory Information Management System (LIMS) combining sample tracking and experiment reporting. %The link between the user and their visits enables data management such as archiving data to a long-term tape storage system, or the Laboratory Information Management System ISPyB (REF. Delageniere et al Bioinformatics 2011). In contrast to the CNB setup, Diamond has integrated the microscopes in much the same fashion to that of the beamlines at the facility, as is shown in figure X

### 2018-04-25 3.workflow_scilifelab.tex
final hopefully

 The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
%Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras? in the same microscope or in different ones, this is confusing, comment when ] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server. %, in order to avoid interruption of data collection due to network issues.

To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III. Hence an script to continuously move the produced data from the adequate storage to the staging server has been created.

>>>>

% The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
%Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras? in the same microscope or in different ones, this is confusing, comment when ] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server. %, in order to avoid interruption of data collection due to network issues.

To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III. Hence a script to continuously move the produced data from the adequate storage to the staging server has been created.

### 2018-04-25 scipionbox.tex
final hopefully

\begin{document}
\title{Using Scipion for Stream Image Processing at Cryo-EM Facilities}
Authors to be sorted later
\author{C.O.S. Sorzano$^{(b)}$,
A. Jim\'{e}nez$^{(b)}$,
R. Melero$^{(b)}$,\\
M. Mart\'{i}nez$^{(b)}$,
J. G\'{o}mez-Blanco$^{(b)}$,
L. del Cano$^{(b)}$,\\
T. Majtner$^{(b)}$,
E. Ram\'{i}rez$^{(b)}$,
P. Conesa$^{(b)}$,\\
J.L. Vilas$^{(b)}$,
R. Marabini$^{(a)}$,
Y. Rancel$^{(b)}$,
J.M. Carazo$^{(b)}$,\\
J.M. de la Rosa-Trev\'{i}n$^{(c)}$,
M. Carroni$^{(c)}$,
S. Fleischmann $^{(c)}$,\\
E. Lindahl$^{(c,d)}$
A.W. Ashton$^{(e)}$,
M. Basham$^{(e)}$,\\
D.K. Clare$^{(e)}$,
K. Savage$^{(e)}$,
C.A. Siebert$^{(e)}$,\\
D. Maluenda$^{(a)}$,\\
Grigory G. Sharov$^{(f)}$
        }

@@

\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(b)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(c)}$ Department of Biochemistry and Biophysics, Science for Life Laboratory, Stockholm University, Stockholm, Sweden}} 
\lyxaddress{{\small$^{(d)}$ Swedish e-Science Research Center, KTH Royal Institute of Technology, Stockholm, Sweden}}
\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, United Kingdom}}
\lyxaddress{{\small$^{(f)}$ MRC Laboratory of Molecular Biology - Structural studies, Cambridge, England, United Kingdom}}

%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm). Figures/tables/plots/etc are located at the document's end. 
\setcounter{figure}{2}%I do not know why but figure counter starts with -1
\textbf{Corresponding Author: Roberto Marabini, Ph = +34 91 5854510, email: roberto@cnb.csic.es\\} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
%Escuela Polit\'{e}cnica Superior\\
%Universidad Aut\'{o}noma de Madrid\\
%28049 Cantoblanco\\
%Madrid\\
%Spain

>>>>

\begin{document}
\title{Using Scipion for Stream Image Processing at Cryo-EM Facilities}

\author{
J. G\'{o}mez-Blanco$^{(a, g)}$,
J.M. de la Rosa-Trev\'{i}n$^{(c, g)}$,
R. Marabini$^{(b, g)}$,\\
L. del Cano$^{(a)}$,
A. Jim\'{e}nez$^{(a)}$,
M. Mart\'{i}nez$^{(a)}$,\\
R. Melero$^{(a)}$,
T. Majtner$^{(a)}$,
D. Maluenda$^{(a)}$,
J. Mota$^{(a)}$,\\
Y. Rancel$^{(a)}$,
E. Ram\'{i}rez$^{(a)}$,
J.L. Vilas$^{(a)}$,\\
M. Carroni$^{(c)}$,
S. Fleischmann $^{(c)}$,
E. Lindahl$^{(c,d)}$\\
A.W. Ashton$^{(e)}$,
M. Basham$^{(e)}$,
D.K. Clare$^{(e)}$,\\
K. Savage$^{(e)}$,
C.A. Siebert$^{(e)}$,\\
Grigory G. Sharov$^{(f)}$,
C.O.S. Sorzano$^{(a)}$,
P. Conesa$^{(a)}$,
J.M. Carazo$^{(a)}$,\\
}

@@

\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(a)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(c)}$ Department of Biochemistry and Biophysics, Science for Life Laboratory, Stockholm University, Stockholm, Sweden}} 
\lyxaddress{{\small$^{(d)}$ Swedish e-Science Research Center, KTH Royal Institute of Technology, Stockholm, Sweden}}
\lyxaddress{{\small$^{(e)}$ Diamond Light Source, Harwell Science and Innovation Campus, Didcot OX11 0DE, United Kingdom}}
\lyxaddress{{\small$^{(f)}$ Medical Research Council Laboratory of Molecular Biology, Francis Crick Avenue, Cambridge CB2 OQH, UK}}
\lyxaddress{{\small$^{(g)}$ These authors  contributed equally to this work.}}

%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm). Figures/tables/plots/etc are located at the document's end. 
\setcounter{figure}{2}%I do not know why but figure counter starts with -1
\textbf{Corresponding Author: Roberto Marabini, Phone = +34 91 5854510, email: roberto@cnb.csic.es\\} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
%Escuela Polit\'{e}cnica Superior\\
%Universidad Aut\'{o}noma de Madrid\\
%28049 Cantoblanco\\
%Madrid\\
%Spain
